The Rockerverse: Packages and Applications for Containerisation with R

### 2020-03-03 manuscript.Rmd
remove rkt, which just got stopped
https://github.com/rkt/rkt/issues/4024

New features which make complex workflows accessible and reproducible and the variety in packages connected with containerisation, even when they have overlapping features, are a signal and a support for a growing user base.
This growth is possibly the most important goal for the foreseeable future in the _Rockerverse_, and, just like the Rocker images have matured over years of use and millions of runs, the new ideas and prototypes will have to prove themselves.
It should be noted that the dominant position is that Docker is a blessing and a curse for these goals.
It might be wise to start experimenting with non-Docker containerisation tools now, e.g., R packages interfacing with other container engines such as [podman/buildah](https://github.com/containers/libpod) or [rkt](https://coreos.com/rkt/), or an R package for creating `Singularity` files.
Such efforts might help to avoid lock-in and to design sustainable workflows based on concepts of _containerisation_, not on their implementation in Docker.
If adoption of containerisation and R continue to grow, the missing pieces for a success predominantly lie in (a) coordination and documentation of activities to reduce repeated work in favour of open collaboration, (b) the sharing of lessons learned from use cases to build common knowledge and language, and (c) a sustainable continuation and funding for development, community support, and education.
A first concrete effort to work towards these missing pieces should be sustain the structure and captured status quo from this work in form of a _CRAN Task View on containerisation_.

>>>>

New features which make complex workflows accessible and reproducible and the variety in packages connected with containerisation, even when they have overlapping features, are a signal and a support for a growing user base.
This growth is possibly the most important goal for the foreseeable future in the _Rockerverse_, and, just like the Rocker images have matured over years of use and millions of runs, the new ideas and prototypes will have to prove themselves.
It should be noted that the dominant position is that Docker is a blessing and a curse for these goals.
It might be wise to start experimenting with non-Docker containerisation tools now, e.g., R packages interfacing with other container engines, such as [podman/buildah](https://github.com/containers/libpod), or an R package for creating `Singularity` files.
Such efforts might help to avoid lock-in and to design sustainable workflows based on concepts of _containerisation_, not on their implementation in Docker.
If adoption of containerisation and R continue to grow, the missing pieces for a success predominantly lie in (a) coordination and documentation of activities to reduce repeated work in favour of open collaboration, (b) the sharing of lessons learned from use cases to build common knowledge and language, and (c) a sustainable continuation and funding for development, community support, and education.
A first concrete effort to work towards these missing pieces should be sustain the structure and captured status quo from this work in form of a _CRAN Task View on containerisation_.
